---
title: "Reproducible research, project #1"
author: "Karl Kuder"
date: "June 30, 2016"
output: html_document
---

# File processing and summaries
Download, unzip and read the input data file.
Convert date from a factor to date format.
Sort the data by date then interval.
Build daily and interval summmaries.

```{r}
download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",
              destfile = "activity.zip")
unzip("activity.zip")
Activity <- read.csv("activity.csv")
Activity$date <- as.Date(Activity$date)
Activity <- Activity[order(Activity$date, Activity$interval),]
DailySum <- aggregate(steps ~ date, data = Activity, na.rm = TRUE, FUN = sum)
IntMean <- aggregate(steps ~ interval, data = Activity,
                     na.rm = TRUE, FUN = mean)

```





# Total number of steps taken each day (original data)
```{r}
hist(DailySum$steps, xlab = "Steps", ylab = "Days", 
     main = "Number of days at each level of steps")
meanSteps <- mean(DailySum$steps)
meanSteps
medianSteps <- median(DailySum$steps)
medianSteps
```

The mean number of steps taken each day is `r format(meanSteps, nsmall=2)`

The median number of steps taken each day is `r medianSteps`



# Daily pattern
```{r}
with(IntMean, plot(interval, steps, type = "l"))
maxInterval <- with(IntMean, interval[(steps == max(steps))])
```

The interval with the maximum average number of steps is `r maxInterval`



# Imputing missing values

Find number of intervals which are missing step data.

```{r}
sum(is.na(Activity$steps))
```

Create a new data set, then loop throuhg it and impute values when steps
are missing by using the average steps for that interval overall.
```{r}
ActivityNew <- Activity
for (i in 1:nrow(ActivityNew)) {
    if (is.na(ActivityNew$steps[i])) {
        ActivityNew$stepSource[i] <- "imputed"
        ActivityNew$newSteps[i] <- 
            IntMean$steps[IntMean$interval == ActivityNew$interval[i]]
    } else {
        ActivityNew$stepSource[i] <- "actual"
        ActivityNew$newSteps[i] <- ActivityNew$steps[i]
    }
}

ActivityNew$stepSource <- as.factor(ActivityNew$stepSource)
```



# Total number of steps taken each day (with imputed data)
```{r}
DailySumNew <- aggregate(newSteps ~ date, data = ActivityNew, na.rm = TRUE, FUN = sum)

hist(DailySumNew$newSteps, xlab = "Steps", ylab = "Days", 
     main = "Number of days at each level of steps")
meanStepsNew <- mean(DailySumNew$newSteps)
meanStepsNew
medianStepsNew <- median(DailySumNew$newSteps)
medianStepsNew
```

Diference in mean and median between original data and new data with 
imputed values for missing data.
```{r}
meanDif <- meanStepsNew - meanSteps
meanDif
medianDif <- medianStepsNew - medianSteps
medianDif
```

The mean is the same, which makes sense since imputed data is based on interval
means. The median daily steps is slightly higher.

Create a new factor variable to classify as weekends or weekdays
```{r}
ActivityNew$dayType <- 
    ifelse(((weekdays(ActivityNew$date)) %in% c('Sunday', 'Saturday')),
           'weekend', 'weekday')
```

